Economic Theory of Fertility Explains Modern Natural Selection

### 2020-12-01 why-negative-selection.Rmd
Add townsend deprivation regressions for parents

n_in_regs <- function (var, dv = "n_children") {
  fh_subset <- famhist
  if (dv == "n_children") fh_subset %<>% filter(kids_ss)
  # all PGS have the same NA pattern, so we just use one
  tbl <- table(fh_subset[
                  ! is.na(fh_subset[dv]) & ! is.na(fh_subset$whr_combined), 
                  var
                ])

@@

add_n <- function (var, famhist_var = var, dv = "n_children") {
  var <- as.factor(var)
  n <- n_in_regs(famhist_var, dv)
  levels(var) <- sprintf("%s (N = %s)", levels(var), 
                         prettyNum(n[seq_len(nlevels(var))], big.mark = ","))
  # var_n <- sprintf("%s (N = %s)", var, prettyNum(n[var], big.mark = ","))
  # var_n[is.na(var)] <- NA
  var
}

@@

@@

Among the parents' generation, we can control for age at first live
birth using the subsets of respondents who reported their mother's or
father's age, and who had no elder siblings. We run regressions on
*number of siblings* on these subsets, controlling for either parent's
age at their birth. Figure \@ref(fig:plot-age-birth-parents) shows the
results. Effect sizes are very similar, whether controlling for father's
or mother's age. Unlike for the respondents' generation, effect sizes
are positively correlated with the effect sizes from bivariate
regressions without the age at birth control (father's age at birth:
$\rho$ `r cor_raw_fath`; mother's age at birth: $\rho$
`r cor_raw_moth`).

>>>>

n_in_regs <- function (var, dv = "n_children", data = famhist) {
  fh_subset <- data
  if (dv == "n_children") fh_subset %<>% filter(kids_ss)
  # all PGS have the same NA pattern, so we just use one
  tbl <- table(fh_subset[
                  ! is.na(fh_subset[dv]) & ! is.na(fh_subset$whr_combined), 
                  var
                ])

@@

add_n <- function (var, famhist_var = var, dv = "n_children", data = famhist) {
  var <- as.factor(var)
  n <- n_in_regs(famhist_var, dv, data)
  levels(var) <- sprintf("%s (N = %s)", levels(var), 
                         prettyNum(n[seq_len(nlevels(var))], big.mark = ","))
  # var_n <- sprintf("%s (N = %s)", var, prettyNum(n[var], big.mark = ","))
  # var_n[is.na(var)] <- NA
  var
}

@@

```{r plot-siblings-townsend, fig.cap = "Selection effects in the parents' generation by Townsend deprivation score quintile", fig.align = "center"}

drake::loadd(res_townsend_parents)
drake::loadd(famhist_townsend_71)
n_regs <- nrow(res_townsend_parents)

res_townsend_parents %>% 
  mutate(
    `Townsend quintile` = sub("Quin71(\\d):.*", "\\1", term),
    add_n(`Townsend quintile`, "Quin71", dv = "n_sibs", data = famhist_townsend_71)
  ) %>% 
  standard_ggplot(`Townsend quintile`, n_regs = n_regs, score_col = score_name, order_idx = 5)

rm(famhist_townsend_71)
```

We also check whether selection effects differ by socio-economic status in the
parents' generation. We have no information about parents' income, so we use
the 1971 Townsend deprivation score of respondents' birthplace as a proxy. Results
(Figure \@ref(fig:plot-siblings-townsend)) show the same pattern as for respondents:
effect sizes are larger and more significant in the most deprived areas.

@@

Among the parents' generation, we can control for age at first live
birth using the subsets of respondents who reported their mother's or
father's age, and who had no elder siblings. We run regressions on
*number of siblings* on these subsets, controlling for either parent's
age at their birth. Figure \@ref(fig:plot-age-birth-parents) shows the
results. Effect sizes are very similar, whether controlling for father's
or mother's age. As in the respondents' generation, effect sizes
are negatively correlated with the effect sizes from bivariate
regressions without the age at birth control, though the correlation
is smaller (father's age at birth: $\rho$ `r cor_raw_fath`; mother's age at
birth: $\rho$ `r cor_raw_moth`).
